---
title: "Chinbook"
output: html_notebook
Description: use data from an SQL database to determine the most popular genre and album based on sales record. Also visualize the findings.
---
1. load libraries

```{r}
#install.packages("RSQLite")
library(RSQLite)
library(DBI)

```
2. make helper functions

```{r}
run_query <- function (query) {
  conn <- dbConnect(SQLite(), "chinook.db")
  result <- dbGetQuery(conn, query)
  dbDisconnect(conn)
  return(result)
}

#start using the combination of single and double quotes
#interesting way of looking at tables as a whole
show_tables <- function () {
query <- 'SELECT
            name,
            type
          FROM sqlite_master
          WHERE type IN ("table","view");'
run_query(query)
}

tables <- show_tables()
```

Quest A. Genre recommendation
A.1, query the absolute and percentage of tracks solde for each Genre
```{r}
query1 <- "
SELECT 
  g.name genre_name,
  COUNT(g.name) track_total,
  ROUND(CAST(COUNT(g.name) AS Float)/
    CAST(
          (
          SELECT 
            COUNT(il.track_id) total_all_genre
          FROM invoice_line il 
          ) AS Float),4) track_percent
FROM genre g
LEFT JOIN track t ON g.genre_id = t.genre_id
LEFT JOIN invoice_line il ON il.track_id = t.track_id
WHERE il.invoice_id IS NOT NULL
GROUP BY g.name
ORDER BY track_total DESC
"

query2 <- "SELECT 
  g.name genre_name,
  COUNT(g.name) track_total,
  ROUND(CAST(COUNT(g.name) AS Float)/
    CAST((SELECT 
            COUNT(il.track_id) total_all_genre
          FROM invoice_line il
          ) AS Float),4) track_percent
FROM invoice_line il
LEFT JOIN track t ON il.track_id = t.track_id
LEFT JOIN genre g ON t.genre_id = g.genre_id
GROUP BY g.name
ORDER BY track_total DESC"

#genre_table1 <- run_query(query1)
genre_table <- run_query(query2)

```

A.2. illustrations
```{r}
genre_in_quest <- c("Hip Hop","Punk","Pop","Blues")
library(dplyr)
library(stringr)
library(ggplot2)
plot_table <- genre_table %>%
  filter(str_detect(genre_name,genre_in_quest[1]) 
                    | str_detect(genre_name,genre_in_quest[2])
                    | str_detect(genre_name,genre_in_quest[3]) 
                    | str_detect(genre_name,genre_in_quest[4]))
ggplot(data = plot_table) + 
  aes(x = genre_name, y = track_total)+ 
  geom_bar(stat = "identity")
ggplot(data = plot_table) + 
  aes(x = genre_name, y = track_percent*100)+ 
  geom_bar(stat = "identity")
```
A.3. Recommendations:

From the sales database, we calculated the number of tracks sold, as well as the percentage of tracks sold for each genre. The No.1 genre that has the most sale is "Rock", whic occupies 55% of the total sale. 

When evaluating the four potential genres, we plotted their total sales and percentage. From the highest to the lowest sales, the four genres follows the order of "Punk", "Blues", "Pop" and "Hip-Hop" with number of tracks sold: 492, 124, 63, 33, and the percentages: 10.3%, 2.6%, 1.3%, 0.7%.

So our recommendations of picking three genres out of the four would be: "Punk", "Blues" and "Pop".

We would also recommending signing more "Rock" genres.

Quest B. Sales support agents performance and potential relating factors
B.1 prepare data for sales support agent
```{r}
query <- "
SELECT 
  e.first_name || ' ' || e.last_name employee_name,
  SUM(i.total) total_sales_amount,
  e.title,
  e.reports_to,
  e.birthdate,
  e.hire_date
FROM employee e
LEFT JOIN customer c ON e.employee_id = c.support_rep_id
LEFT JOIN invoice i ON c.customer_id = i.customer_id
GROUP BY employee_name
ORDER BY total_sales_amount DESC
"
employee_table <- run_query(query)

employee_table_organized <- employee_table %>%
  filter(!is.na(total_sales_amount))%>%
           mutate(hire_date = as.Date(hire_date),birthdate = as.Date(birthdate)) %>%
  mutate(hire_age = hire_date - birthdate) %>%
  mutate(hire_age_year = round(as.numeric(hire_age)/365.25)) %>%
  mutate(sex = c("Female","Female","Male")) %>%
  select(employee_name, total_sales_amount, hire_date, hire_age_year, sex)
```
B.2. illustration
```{r}
ggplot(data = employee_table_organized) + 
  aes(x = employee_name, y = total_sales_amount, size = hire_date, color = hire_age_year, shape = sex)+
  geom_point()
```

B.3 Summary

The three sales support agents ranked from top to bottom sales as follows: Jane Peacock, Margaret Park, and steven Johnson. There are three potential contibutors to the amount of total sales.

1. Age at hires-- younger agent produces more sales
The top 1 agent was the youngest at hire 44 yrs (the darker color of the shape shows the younger age), and the other two were 50 and older. Younger agent could be more advantageous in communicating with customers and thus potentially boost the total sales.

2. Sex of the sales person-- female agent produces more sales
The top 1 and 2 agents are both females (circles in the plot) and the bottom one is a male (triangle). There could be a sex difference in how agents interact with customers and could potentially contribute to the difference in total amount of sales.

3. Duration at job -- an agent had longer duration at the job produces more sales than someone who had shorter duration at the job
The top 1 and 2 agents were hired earlier than the third sales person, (the time of hire represented by the size of the shape, with earlier hire having a smaller size than the later hire,) thus would have worked at the job longer. Longer time at job would accumulate more experience and could potentially contribute to a higher sale amount.

Quest C. collates data on purchases from different countries

total number of customers
total value of sales
average value of sales per customer
average order value

C.1 create the query
```{r}
query <- '
WITH 
  customer_info AS (
SELECT
  country,
  COUNT(DISTINCT(customer_id)) num_customer,
  CASE
    WHEN COUNT(DISTINCT(customer_id)) =1 THEN "Other"
    ELSE country
    END AS country_category
FROM customer c
GROUP BY country
),
  customer_info2 AS(
SELECT 
  country,
  country_category,
  SUM(num_customer) num_customer
FROM customer_info
GROUP BY country_category
  )
SELECT
  c.country,
  cc.country_category,
  cc.num_customer,
  SUM(i.total) total_sale,
   CAST(SUM(i.total) AS Float)/CAST(cc.num_customer AS Float) sale_per_customer,
   CAST(SUM(i.total) AS Float)/CAST(COUNT(i.invoice_id) AS Float) sale_per_invoice
FROM customer_info2 cc
LEFT JOIN customer c ON cc.country = c.country
LEFT JOIN invoice i ON c.customer_id = i.customer_id
GROUP BY country_category
ORDER BY total_sale DESC
'
country_table <- run_query(query)
```
C.2 visulization
```{r}
#total number of customers
plot_country_table <- country_table %>%
  filter(country_category != "Other")
ggplot(data = plot_country_table) + 
  aes(x = country, y = num_customer) +
  geom_bar(stat = "identity")

ggplot(data = plot_country_table) + 
  aes(x = country, y = total_sale) +
  geom_bar(stat = "identity")

ggplot(data = plot_country_table) + 
  aes(x = country, y = sale_per_customer) +
  geom_bar(stat = "identity")

ggplot(data = plot_country_table) + 
  aes(x = country, y = sale_per_invoice) +
  geom_bar(stat = "identity")
```
C.3 recommendations:
In general having more customers and increase sales per customer would boost the total sale for all countries. 

From all the countries, USA has the largest amount of total sale, and the most number of customers. But the sale amount per customer and per invoice, USA is similar to most of the other countries. It means similar strategy should be used on USA to other countries, meaning increasing the number of customers and increasing the amout of sales per customer.

Czech Republic had the highest sale per customer, but it only has a small number of customers. Increasing the number of customers in Czech Republic would be a way to most likely bring boost of total sale.

Question D. evaluate the sales of whole album versus combination of tracks
D.1. prepare the data
why not just take all the tables and use dataframe to analyze them? Would be much simpler!

```{r}
query_template <- function(invoice_id) {
  query <- 
str_c('
WITH 
  invoice_to_album AS(
SELECT 
  i.invoice_id,
  AVG(t.album_id) one_album_id
FROM invoice i
LEFT JOIN invoice_line il ON i.invoice_id = il.invoice_id
LEFT JOIN track t ON il.track_id = t.track_id
GROUP BY i.invoice_id
),

  invoice_to_track AS(
SELECT 
  i.invoice_id,
  il.track_id
  FROM invoice i
LEFT JOIN invoice_line il ON i.invoice_id = il.invoice_id
WHERE i.invoice_id == ', invoice_id, '
),

  album_to_track AS(
SELECT
  album_id,
  track_id
FROM track
WHERE album_id == (
SELECT one_album_id FROM invoice_to_album
WHERE invoice_id == ', invoice_id, ')
)

SELECT
invoice_id,
CASE
  WHEN
    (SELECT track_id FROM invoice_to_track
      EXCEPT
    SELECT track_id FROM album_to_track
    ) IS NULL
    AND
    (SELECT track_id FROM invoice_to_track
    EXCEPT
    SELECT track_id FROM album_to_track
    ) IS NULL THEN "1"
  ELSE "0"
  END AS whole_album
FROM invoice
WHERE invoice_id == ', invoice_id
  )
return(query)
}

query1 <- '
SELECT invoice_id,
invoice_date,
total
FROM invoice
ORDER BY invoice_id
'
invoice_table <- run_query(query1)
invoice_list <- invoice_table$invoice_id

for (invoice_id in invoice_list) {
  print(invoice_id)
  query <- query_template(invoice_id)
  if (invoice_id == invoice_list[1]) {
    invoice_category <- run_query(query)
  }else {
    invoice_category <- bind_rows(invoice_category,run_query(query))
  }
}
invoice_table<- invoice_table %>% left_join(invoice_category, on = "invoice_id")
```

D.2 illustration
```{r}
ggplot(data = invoice_table) + 
  aes(x = whole_album) + 
  geom_bar()

ggplot(data = invoice_table) + 
  aes(x = whole_album, y = total) + 
  geom_bar(stat = "identity")

ggplot(data = invoice_table) + 
  aes(x = whole_album, y = total) + 
  geom_boxplot()

ggplot(data = invoice_table) + 
  aes(x = invoice_date, y = total, color = whole_album) + 
  geom_point()

```

D.3 recommendation
Comparing the two ways of selling tracks, 1 as a whole album, and 2 as manually picked tracks, method 2 had twice as many sales as method 1 (figure 1). Method 2 also generated 1.5 times total sales amount as method 1 (figure 2). Method 1 has a larger amount per sale than metho 2 (figure 3 and 4). It showed that selling music by tracks would create more total sales and total revenue, thus is recommended.

